Distributional Trees and Forests

### 2017-10-09 pkg/disttree/inst/draft/draft.Rnw
further descriptions in the rain section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1644 255f4249-2263-4417-b1ea-8498779e27a7

\begin{tabular}{| l | c   c   c   c |}
\hline
 & \textbf{mu link} & \textbf{sigma link} & \textbf{mu covariates} & \textbf{sigma covariates} \\
\hline
\textbf{disttree} & identity & log & all covariates & all covariates \\ 
\hline
\textbf{distforest} & identity & log & all covariates & all covariates \\
\hline
\textbf{gamlss} & identity & log & tppow{\_}mean, & tppow{\_}sprd,\\
& & & tppow{\_}max, & dswrf{\_}sprd{\_}mean, \\
& & & tppow{\_}mean1218, & tppow{\_}sprd1218,\\ 
& & & capepow{\_}mean1218, & capepow{\_}mean1218,\\
& & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
& & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
& & & pwat{\_}mean{\_}mean, & \\
& & & tdiff500850{\_}mean, & \\
& & & msl{\_}diff & \\
\hline
\textbf{gamboostLSS} & identity & log & all covariates & all covariates \\
\hline
\textbf{EMOS} & identity & log & tppow{\_}mean & tppow{\_}spread \\
\hline
\end{tabular}
\\

manuall selected covariates in gamlss\\
mention interactions:\\
tppow{\_}mean1218 $\cdot$ capepow{\_}mean1218,\\
tppow{\_}sprd1218 $\cdot$ capepow{\_}mean1218\\
\\
(advantage: no sepcification needed in disttree/distforest)

\SweaveOpts{eval = TRUE}
\newpage
\subsection{Station Axams}
one station \\
Axams, closest to Ibk\\
865 observations\\

\subsubsection{Predictions}
Fit one model on the first 25 years (1985--2009, both included)\\
predictions for the 26th, 27th and 28th year\\
plot predicted distributions together with obersvations for one day in July of 26th, 27th and 28th year  
<<echo=FALSE, results=hide>>=
#### Axams prediction 25 - 3
if(file.exists("rain_Axams_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred.rda")
  load("rain_Axams_pred.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred.R")
  source("rain_axams_pred.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = TRUE,
                         frac = FALSE)

@@

Variable importance\\
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp100_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  load("vimp100_crps.rda")
} else {

>>>>

\begin{tabular}{| l | c   c   c   c |}
\hline
 & \textbf{mu link} & \textbf{sigma link} & \textbf{mu covariates} & \textbf{sigma covariates} \\
\hline
\textbf{disttree} & identity & log & all covariates & all covariates \\ 
\hline
\textbf{distforest} & identity & log & all covariates & all covariates \\
\hline
\textbf{gamlss} & identity & log & tppow{\_}mean, & tppow{\_}sprd,\\
& & & tppow{\_}max, & dswrf{\_}sprd{\_}mean, \\
& & & tppow{\_}mean1218, & tppow{\_}sprd1218,\\ 
& & & capepow{\_}mean1218, & capepow{\_}mean1218,\\
& & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
& & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
& & & pwat{\_}mean{\_}mean, & \\
& & & tdiff500850{\_}mean, & \\
& & & msl{\_}diff & \\
\hline
\textbf{gamboostLSS} & identity & log & all covariates & all covariates \\
\hline
\textbf{EMOS} & identity & log & tppow{\_}mean & tppow{\_}spread \\
\hline
\end{tabular}
\\
For the gamlss model the covariates were selected manually based on expert knowledge. Additionally, interactions are specified in advance, once between tppow{\_}mean1218 and capepow{\_}mean1218 for the location parameter and once between tppow{\_}sprd1218 and capepow{\_}mean1218 for the l ocatioin parameter. By putting extra effort in the specification of the formulas the gamlss model is provided with additional information and is therefore expected to profit from this advantage.\\

\SweaveOpts{eval = TRUE}
\newpage
\subsection{Station Axams}
First of all we focus on one observation station, Axams in Tyrol. This station is selected due to its geographical closeness to Innsbruck, the capital of Tyrol.\\
As for all other stations, daily observations of the month of July of 28 years are available, starting in 1985 and ending in 2012.\\

\subsubsection{Predictions}
In a first step one model is fit on the first 25 years (1985--2009, both included) and then predictions for the 26th, 27th and 28th year (2010, 2011 and 2012) are made.\\
To illustrate the results, one day of July is chosen and the estimated distributions for this day of all three prediction years are plotted together with the corresponding obersvations.  
<<echo=FALSE, results=hide>>=
#### Axams prediction 25 - 3
if(file.exists("rain_Axams_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred.rda")
  load("rain_Axams_pred.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred.R")
  source("rain_axams_pred.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = TRUE,
                         frac = FALSE)

@@

Analyzing variable importance of the distforest model leads to the following list of the covariates which show the highest values in variable importance.\\
In the following plot the marginal effect of these variables on the distribution parameters are plotted. (All other covariates are kept at their mean.)
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp100_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp100_crps.rda")
  load("vimp100_crps.rda")
} else {
